---
title: "Distribution Map: Bombus and its Subgenera"
output:
  word_document: default
  pdf_document: default
  html_notebook: default


```{r}
library('rgdal')
library('dismo')
library('raster')
library('sp')
library('maptools')
library('readr')
library('ggplot2')
library('ggmap')
library('sf')

```


```{r}
#saving file into a label and reading
occ.data <- read.csv(file = "data/bombus_emu.csv")

summary(occ.data)
```

```{r}
dir.create(path = 'data')
dir.create(path = 'output')

```


```{r}
#shows properties of data, rows and columns
dim(occ.data)
#Column names
colnames(occ.data)

#bioclimatic data
bioclim.data<- getData(name = 'worldclim',
                       var = 'bio',
                       res = 2.5,
                       path = "data/")
```


```{r}
#disregarding all nulled values for lon and lat
clean.data <- subset(occ.data, !is.na(DarDecimalLatitude) & !is.na(DarDecimalLongitude))


```
```{r}

dim(clean.data)
clean.data

```






```{r}
#Determining what the max and min geographical values of map are
max.lat <- (max(clean.data$DarDecimalLatitude))
min.lat <- (min(clean.data$DarDecimalLatitude))
max.lon <- (max(clean.data$DarDecimalLongitude))
min.lon <- (min(clean.data$DarDecimalLongitude))
```
```{r}
#Display Values
max.lat
min.lon
max.lon
min.lat



```
```{r}
#Establishing extent of map (will be able to plot only north american in later model)
geo.extent <- extent(x = c(min.lon, max.lon, min.lat, max.lat))
c(min.lon, max.lon, min.lat, max.lat)
```


```{r}
#Using only collumns from lon and lat
loc.data <- clean.data [, c('DarDecimalLongitude', 'DarDecimalLatitude')]
loc.data

```

```{r}
data("wrld_simpl")

plot(wrld_simpl,
     xlim = c(min.lon, max.lon),
     ylim = c(min.lat, max.lat ), 
     axes = TRUE,
     col = 'grey95')

#adding points on to map
points(x = loc.data$DarDecimalLongitude,
       y = loc.data$DarDecimalLatitude,
       col = 'olivedrab',
       pch = 20,
       cex = 0.75)
#box around map
box()
```


```{r}
bc.model <- bioclim(x = bioclim.data, p = loc.data)

predict.presence <- dismo::predict(object = bc.model, x = bioclim.data, ext = geo.extent)
```
```{r}
#plot base map
plot(wrld_simpl,
     xlim = c(min.lon, max.lon),
     ylim = c(min.lat, max.lat),
     axes = TRUE,
     col = "grey95")

#adding model probabilities
plot(predict.presence, add = TRUE)
 
#redrawing country borders
plot(wrld_simpl, add = TRUE, border = "grey6")

#adding OG observations
points(x = loc.data$DarDecimalLongitude,
       y = loc.data$DarDecimalLatitude,
       col = 'olivedrab',
       pch = 20,
       cex = 0.75)
```
```{r}
sep.data <- clean.data [, c('Taxon', 'DarDecimalLongitude', 'DarDecimalLatitude')]

```
```{r}
pyro.data <- read.csv(file = 'data/Pyrobombus_emu.csv')

locpyro.data <- pyro.data [, c('DarDecimalLongitude', 'DarDecimalLatitude')]

locpyro.data
```
```{r}
plot(wrld_simpl,
     xlim = c(min.lon, max.lon),
     ylim = c(min.lat, max.lat),
     axes = TRUE,
     col = "grey95")

plot(wrld_simpl, add = TRUE, border = "grey5")

points(x = clean.data$DarDecimalLongitude,
       y = clean.data$DarDecimalLatitude,
       col = 'olivedrab',
       pch = 20,
       cex = 0.75)

points(x = locpyro.data$DarDecimalLongitude,
       y = locpyro.data$DarDecimalLatitude,
       col = 'red',
       pch = 20,
       cex = 0.75)
```
```{r}
Agro.data <- read.csv(file = 'data/Agro_emu.csv')
Alpigeno.data <- read.csv(file = 'data/Alpigeno_emu.csv')
Alpino.data <- read.csv(file = 'data/Alpino_emu.csv')
Bombias.data <- read.csv(file = 'data/Bombias_emu.csv')
BombusSub.data <- read.csv(file = 'data/BombusSub_emu.csv')
Brachy.data <- read.csv (file = 'data/Brachycephalobombus_emu.csv')
Coccineo.data <- read.csv(file = 'data/Coccineo_emu.csv')
Confusii.data <- read.csv(file = 'data/Confusii_emu.csv')
Crotchii.data <- read.csv(file = 'data/Crotchii_emu.csv')
Collumano.data <- read.csv(file = 'data/Collumano_emu.csv')
Diverso.data <- read.csv(file = 'data/Diverso_emu.csv')
Dasy.data <- read.csv(file = 'data/Dasy_emu.csv')
Fervido.data <- read.csv(file = 'data/Fervido_emu.csv')
Thoraco.data <- read.csv(file = 'data/Thoraco_emu.csv')
```
```{r}
#Disregarding all na values in each data frame
Alpigeno.data <- subset(Alpigeno.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locAlpigen.data <- Alpigeno.data [, c('DarLongitude', 'DarLatitude')]
locAlpigen.data

Agro.data <- subset(Agro.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locAgro.data <- Agro.data [, c('DarLongitude', 'DarLatitude')]

Alpino.data <- subset(Alpino.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locAlpino.data <- Alpino.data [, c('DarLongitude', 'DarLatitude')]

Bombias.data <- subset(Bombias.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locBombias.data <- Bombias.data [, c('DarLongitude', 'DarLatitude')]


BombusSub.data <- subset(BombusSub.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locBombus.data <- BombusSub.data [, c('DarLongitude', 'DarLatitude')]


Brachy.data <- subset(Brachy.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locBrachy.data <- Brachy.data [, c('DarLongitude', 'DarLatitude')]

#skipped Coccineo

Collum.data <- subset(Collumano.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locCollum.data <- Collum.data [, c('DarLongitude', 'DarLatitude')]

Thoraco.data <- subset(Thoraco.data, !is.na(DarLatitude) & !is.na(DarLongitude))
locThoraco.data <- Thoraco.data [, c('DarLongitude', 'DarLatitude')]
```
```{r}
plot(wrld_simpl,
     xlim = c(min.lon, max.lon),
     ylim = c(min.lat, max.lat),
     axes = TRUE,
     col = "grey95")

plot(wrld_simpl, add = TRUE, border = "grey5")

#points(x = clean.data$DarDecimalLongitude,
#      y = clean.data$DarDecimalLatitude,
#     col = 'olivedrab',
#    pch = 20,
#   cex = 0.75)

points(x = locpyro.data$DarDecimalLongitude,
       y = locpyro.data$DarDecimalLatitude,
       col = 'red',
       pch = 20,
       cex = 0.75)

points(x = locAlpigen.data$DarLongitude,
       y = locAlpigen.data$DarLatitude,
       col = 'yellow',
       pch = 20,
       cex = 0.75)

points(x = locAgro.data$DarLongitude,
       y = locAgro.data$DarLatitude,
       col = 'pink',
       pch = 20,
       cex = 0.75)
 
points(x = locAlpino.data$DarLongitude,
       y = locAlpino.data$DarLatitude,
       col = 'blue',
       pch = 20,
       cex = 0.75)


points(x = locBombias.data$DarLongitude,
       y = locBombias.data$DarLatitude,
       col = 'purple',
       pch = 20,
       cex = 0.75)

points(x = locBombus.data$DarLongitude,
       y = locBombus.data$DarLatitude,
       col = 'grey',
       pch = 20,
       cex = 0.75)

points(x = locBrachy.data$DarLongitude,
       y = locBrachy.data$DarLatitude,
       col = 'olivedrab',
       pch = 20,
       cex = 0.75)

points(x = locCollum.data$DarLongitude,
       y = locCollum.data$DarLatitude,
       col = 'black',
       pch = 20,
       cex = 0.75)

points(x = locThoraco.data$DarLongitude,
       y = locThoraco.data$DarLatitude,
       col = 'green',
       pch = 20,
       cex = 0.75)
```
```{r}
plot(wrld_simpl,
     xlim = c(-170, -39),
     ylim = c(11, 72),
     axes = TRUE,
     col = "grey95")
points(x = locpyro.data$DarDecimalLongitude,
       y = locpyro.data$DarDecimalLatitude,
       col = 'red',
       pch = 20,
       cex = 0.75)

points(x = locBombus.data$DarLongitude,
       y = locBombus.data$DarLatitude,
       col = 'grey',
       pch = 20,
       cex = 0.75)
points(x = locCollum.data$DarLongitude,
       y = locCollum.data$DarLatitude,
       col = 'black',
       pch = 20,
       cex = 0.75)

points(x = locThoraco.data$DarLongitude,
       y = locThoraco.data$DarLatitude,
       col = 'green',
       pch = 20,
       cex = 0.75) 
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

